---
title: "Climate Data Organization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compile data into single data frame

```{r data compilation}
setwd("~/Documents/Grad School/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly")

month_files <- list.files("~/Documents/Grad School/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly",full.names = T) 

month_files_name <- list.files("~/Documents/Grad School/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly",full.names = F)

indices.monthly = data.frame()

######Read each monthly csv and bind all into one data frame with necessary labels
for(i in 1:length(month_files)){
  data.ind <- read.csv(month_files[i],header = F,skip = 1)
  Month <- rep(substr(month_files_name[i],1,3),times=length(data.ind[,1]))
  index <- rep(substr(month_files_name[i],5,nchar(month_files_name[i])-4),times=length(data.ind[,1]))
  binding = cbind(data.ind,Month,index)
  indices.monthly = rbind(indices.monthly,binding)
    }

Years = 1980:(1980+ncol(indices.monthly)-6)
colnames(indices.monthly) = c("Transect","Lat",'Lon',Years,"Month","Index")  

```

Now let's put this in long form to make analysis easier
```{r monthly long form}
library(data.table)
head(indices.monthly)
indices.monthly.long = data.table(melt(indices.monthly,id=c("Transect","Month","Index"),measure=4:(ncol(indices.monthly)-2)))
head(indices.monthly.long)
class(indices.monthly.long)
index.names = as.character(indices.monthly.long[,unique(Index)])
```

#Visualize the monthly indices 

##Index Key:
FD -Annual count of days when TN (daily minimum temperature) < 0°C  
SU - Annual count of days when TX (daily maximum temperature) > 25°C.  
ID - Annual count of days when TX (daily maximum temperature) < 0 °C.  
TR - Annual count of days when TN (daily minimum temperature) > 20 °C.   
GSL - Growing Season length. Annual* count between the first span of at least 6 days with daily mean temperature TG >5 °C and the first span after July 1st (Jan 1st in SH) of 6 days with TG <5 °C.  
TXx - Monthly maximum value of daily maximum temperature  
TNx - Monthly maximum value of daily minimum temperature  
TXn - Monthly minimum value of daily minimum temperature  
TNn - Monthly minimum value of daily minimum temperature  
TN10p - Percentage of days when TN < 10th percentile  
TX10p - Percentage of days when TX < 10th percentile  
TN90p - Percentage of days when TN > 90th percentile  
TX90p - Percentage of days when TX > 90th percentile  
WSDI - Warm spell duration index: annual count of days with at least 6 consecutive days when TX > 90th percentile  
CSDI - Cold spell duration index: annual count of days with at least 6 consecutive days when TN < 10th percentile  
DTR - Daily temperature range  
Rx1day - Monthly maximum 1-day precipitation  
Rx5day - Monthly maximum consecutive 5-day precipitation  
SDII - Simple precipitation intensity index  
R10mm - Annual count of days when PRCP ≥ 10mm  
R20mm - Annual count of days when PRCP ≥ 20mm  
Rnnmm - Annual count of days when PRCP ≥ nn mm, where nn is a user-defined threshold  
CDD - Maximum length of dry spell: maximum number of consecutive days with RR < 1mm  
CWD - Maximum length of wet spell: maximum number of consecutive days with RR ≥ 1mm  
R95p - Annual total PRCP when RR > 95th percentile  
R99p - Annual total PRCP when RR > 99th percentile  
PRCPTOT - Annual total precipitation on wet days  

```{r monthly visualization}
library(ggplot2)
for(t in 1:length(index.names)) {
plots = ggplot(indices.monthly.long[Index==index.names[t]],aes(x=Month,y=value,fill=Month))+
  geom_violin(trim=FALSE) +
  geom_boxplot(width=0.1,fill="white")+
  labs(title="Transect Climate Index Variability by Month") +
  ylab(index.names[t])+
  theme_classic() 
  
print(plots)
  }

```

##Compilation for yearly indices

```{r data compilation 2}
setwd("~/Documents/Grad School/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly")

year_files <- list.files("~/Documents/Grad School/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly",full.names = T) 

year_files_name <- list.files("~/Documents/Grad School/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly",full.names = F)

indices.yearly = data.frame()

######Read each monthly csv and bind all into one data frame with necessary labels
for(i in 1:length(year_files)){
  data.ind <- read.csv(year_files[i],header = F,skip = 1)
  index <- rep(substr(year_files_name[i],1,nchar(year_files_name[i])-4),times=length(data.ind[,1]))
  binding = cbind(data.ind,index)
  indices.yearly = rbind(indices.yearly,binding)
    }

Years = 1980:(1980+ncol(indices.yearly)-5)
colnames(indices.yearly) = c("Transect","Lat",'Lon',Years,"Index")  

```

Now let's put this in long form to make analysis easier
```{r yearly long form}
library(data.table)
head(indices.yearly)
indices.yearly.long = data.table(melt(indices.yearly,id=c("Transect","Index"),measure=4:(ncol(indices.yearly)-2)))
head(indices.yearly.long)
class(indices.yearly.long)
index.names = as.character(indices.yearly.long[,unique(Index)])
```

Visualize the monthly indices 
```{r yearly visualization}
library(ggplot2)
for(t in 1:length(index.names)) {
plots = ggplot(indices.yearly.long[Index==index.names[t]],aes(x=variable,y=value,fill=variable))+
  geom_violin(trim=FALSE) +
  geom_boxplot(width=0.1,fill="white")+
  labs(title="Transect Climate Yearly-Index Variability") +
  ylab(index.names[t])+
  theme_classic() 
  
print(plots)
  }

```

#Export
Combine Monthly and Yearly Indices into one table and export.

```{r Export}
setwd("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis")

library(data.table)

#Save cover file
fwrite(indices.yearly.long,"/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis/Indices_Yearly.csv")

#save master transect file
fwrite(indices.monthly.long,"/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis/Indices_Monthly.csv")


```



