---
title: "Climate Data Organization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(data.table)
```

## Compile data into single data frame

```{r data-compilation}
setwd("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly")

month_files <- list.files("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly",full.names = T) 

month_files_name <- list.files("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly",full.names = F)

indices.monthly = data.frame(Year = 1, Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA, Index = NA)

######Read each monthly csv and bind all into one data frame with necessary labels
for(i in 1:length(month_files)){
  data.ind <- read.csv(month_files[i],header = F,skip = 1)
  index <- rep(substr(month_files_name[i],1,nchar(month_files_name[i])-4),times=length(data.ind[,1]))
  binding = cbind(data.ind,index)
  names(binding) = names(indices.monthly)
  indices.monthly = rbind(indices.monthly,binding)
    }

indices.monthly = as.data.table(indices.monthly)

#remove placeholder row and duplicate year column
indices.monthly = indices.monthly[!1,]
indices.monthly[,Year:=NULL]
setnames(indices.monthly,"Year.1","Year")


```

Now let's put this in long form to make analysis easier
```{r monthly-long-form}
head(indices.monthly)
indices.monthly.long = melt(indices.monthly,id=c("TransectID","Year","Index"),measure=1:(ncol(indices.monthly)-3))
head(indices.monthly.long)
setnames(indices.monthly.long, "variable", 'Month')

index.names = as.character(indices.monthly.long[,unique(Index)])



```

#Visualize the monthly indices 
Moved to other Rmd file for figure construction

##Index Key:
FD -Annual count of days when TN (daily minimum temperature) < 0°C  
SU - Annual count of days when TX (daily maximum temperature) > 25°C.  
ID - Annual count of days when TX (daily maximum temperature) < 0 °C.  
TR - Annual count of days when TN (daily minimum temperature) > 20 °C.   
GSL - Growing Season length. Annual* count between the first span of at least 6 days with daily mean temperature TG >5 °C and the first span after July 1st (Jan 1st in SH) of 6 days with TG <5 °C.  
TXx - Monthly maximum value of daily maximum temperature  
TNx - Monthly maximum value of daily minimum temperature  
TXn - Monthly minimum value of daily minimum temperature  
TNn - Monthly minimum value of daily minimum temperature  
TN10p - Percentage of days when TN < 10th percentile  
TX10p - Percentage of days when TX < 10th percentile  
TN90p - Percentage of days when TN > 90th percentile  
TX90p - Percentage of days when TX > 90th percentile  
WSDI - Warm spell duration index: annual count of days with at least 6 consecutive days when TX > 90th percentile  
CSDI - Cold spell duration index: annual count of days with at least 6 consecutive days when TN < 10th percentile  
DTR - Daily temperature range  
Rx1day - Monthly maximum 1-day precipitation  
Rx5day - Monthly maximum consecutive 5-day precipitation  
SDII - Simple precipitation intensity index  
R10mm - Annual count of days when PRCP ≥ 10mm  
R20mm - Annual count of days when PRCP ≥ 20mm  
Rnnmm - Annual count of days when PRCP ≥ nn mm, where nn is a user-defined threshold  
CDD - Maximum length of dry spell: maximum number of consecutive days with RR < 1mm  
CWD - Maximum length of wet spell: maximum number of consecutive days with RR ≥ 1mm  
R95p - Annual total PRCP when RR > 95th percentile  
R99p - Annual total PRCP when RR > 99th percentile  
PRCPTOT - Annual total precipitation on wet days  


##Compilation for yearly indices

```{r data-compilation-2}
setwd("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly")

year_files <- list.files("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly",full.names = T) 

year_files_name <- list.files("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly",full.names = F)

indices.yearly = data.table(Row = NA, TransectID = NA, 
                            "1980" = 1, "1981" = 1, "1982" = 1, "1983" = 1, "1984" = 1, "1985" = 1, "1986" = 1, "1987" = 1, "1988" = 1, "1989" = 1,
                            "1990" = 1, "1991" = 1, "1992" = 1, "1993" = 1, "1994" = 1, "1995" = 1, "1996" = 1, "1997" = 1, "1998" = 1, "1999" = 1,
                            "2000" = 1, "2001" = 1, "2002" = 1, "2003" = 1, "2004" = 1, "2005" = 1, "2006" = 1, "2007" = 1, "2008" = 1, "2009" = 1,
                            "2010" = 1, "2011" = 1, "2012" = 1, "2013" = 1, "2014" = 1, "2015" = 1, "2016" = 1, "2017" = 1, "2018" = 1,  Index = NA)

######Read each monthly csv and bind all into one data frame with necessary labels
for(i in 1:length(year_files)){
  data.ind <- data.table(read.csv(year_files[i],header = T,skip = 0))
  data.ind[,X2019:=NULL]
  index <- rep(substr(year_files_name[i],1,nchar(year_files_name[i])-4),times=length(data.ind[,1]))
  binding = cbind(data.ind,index)
  names(binding) = names(indices.yearly)
  indices.yearly = rbind(indices.yearly,binding)
    }

indices.yearly = indices.yearly[!1,] #remove starter row
indices.yearly[,Row:=NULL]

```

Now let's put this in long form to make analysis easier
```{r yearly long form}
head(indices.yearly)
indices.yearly.long = melt(indices.yearly,id=c("TransectID","Index"),measure=2:(ncol(indices.yearly)-1))
head(indices.yearly.long)
setnames(indices.yearly.long, "variable", 'Year')
index.names = as.character(indices.yearly.long[,unique(Index)])
```

##Prepping and merging indices

See model indices selection logic for which ones we are using and why. Several of these need to be adjusted, averages, etc. and then all need to be appropriately merged to the cover file.

Seasonal definitions: HUGH - modify this based on Daren and co responses

Note: monthly calcs do not inherently solve the issue monitoring timing vs weather. I will need to either lag all or just some of the months. Using July 1 as the cutoff

```{r indices prep}

#load in cover
#load data
cover=as.data.table(read.csv("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis/Cover_data_model.csv",header = T)) 

#Create a new cover file to merge climate data into
cover.climate = cover

transect.master=as.data.table(read.csv("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis/Transect_master_V2.csv",header = T)) #use V2 after data issue


indices.yearly=as.data.table(indices.yearly.long)
indices.monthly=as.data.table(indices.monthly.long)
setnames(indices.monthly,c("value","Index"),c("val.mth","Index.m"),skip_absent=TRUE)
setnames(indices.yearly,c("value","Index"),c("val.yr","Index.y"),skip_absent=TRUE)

indices.monthly$Year = as.integer(as.character(indices.monthly$Year))
indices.yearly$Year = as.integer(as.character(indices.yearly$Year))

#Create a new monthly table which changes the year value based on month
indices.monthly.mod = indices.monthly
indices.monthly.mod[Month=="Jul" | Month=="Aug" | Month=="Sep" | Month=="Oct" | Month=="Nov" | Month=="Dec",Year:=Year-1]

###txx
txx = indices.monthly.mod[Index.m=="txx" & Year>2008,.SD[which.max(val.mth)],by=.(Year,TransectID)]

#Merge txx with cover
txx[,Month:=NULL]
txx[,Index.m:=NULL]
txx[order(Year),txx.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,txx, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'txx')

###txxx.win
txx.win = indices.monthly.mod[Index.m=="txx" & Year>2008 & (Month=="Dec" | Month=="Jan" | Month=="Feb"),.SD[which.max(val.mth)],by=.(Year,TransectID)]

#Merge txx.win with cover
txx.win[,Month:=NULL]
txx.win[,Index.m:=NULL]
txx.win[order(Year),txx.win.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,txx.win, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'txx.win')

###tn90p
#First sum tn90p values for winter "dormant" months, let's say Decemeber - March
#tn90p = indices.monthly.mod[Index.m=="tn90p" & Year>2008 & (Month =="Dec" | Month == "Jan" | Month == "Feb" | Month=="Mar"),sum(val.mth)/4,by=.(Year,TransectID)]
#setnames(tn90p,"V1","tn90p")
#tn90p[order(Year),tn90p.lag:= shift(tn90p), by =.(TransectID)]

#Merge tn90p with cover
#cover.climate = merge(cover.climate,tn90p, by=c("Year","TransectID"),all.x = T)

###tx90p.win
#First sum tn90p values for winter "dormant" months, let's say Decemeber - March
tx90p = indices.monthly.mod[Index.m=="tx90p" & Year>2008 & (Month =="Dec" | Month == "Jan" | Month == "Feb" | Month=="Mar"),sum(val.mth)/4,by=.(Year,TransectID)]
setnames(tx90p,"V1","tx90p.win")
tx90p[order(Year),tx90p.win.lag:= shift(tx90p.win), by =.(TransectID)]

#Merge tx90p with cover
cover.climate = merge(cover.climate,tx90p, by=c("Year","TransectID"),all.x = T)

###rx1
#I don't see what this will capture that rx5 will not, hold off for now

###rx5
#This is a monthly index, break out by seasons (Winter/Dormant,Spring, Summer/Fall) and also calculate yearly

#rx5.win
#rx5.win = indices.monthly.mod[Index.m=="rx5day" & Year>2008 & (Month =="Dec" | Month == "Jan" | Month == "Feb" | Month=="Mar"),.SD[which.max(val.mth)],by=.(Year,TransectID)]
#rx5.win[,Month:=NULL]
#rx5.win[,Index.m:=NULL]
#rx5.win[order(Year),rx5.win.lag:= shift(val.mth), by =.(TransectID)]

#cover.climate = merge(cover.climate,rx5.win, by=c("Year","TransectID"),all.x = T)
#setnames(cover.climate,"val.mth",'rx5.win')

#rx5.spr
rx5.spr = indices.monthly.mod[Index.m=="rx5day" & Year>2008 & (Month =="Apr" | Month == "May" | Month =="Jun"),.SD[which.max(val.mth)],by=.(Year,TransectID)]
rx5.spr[,Month:=NULL]
rx5.spr[,Index.m:=NULL]
rx5.spr[order(Year),rx5.spr.lag:= shift(val.mth), by =.(TransectID)]

cover.climate = merge(cover.climate,rx5.spr, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'rx5.spr')

#rx5.summer
rx5.sum = indices.monthly.mod[Index.m=="rx5day" & Year>2009 & ( Month == "Jul"| Month == "Aug"),
                             .SD[which.max(val.mth)],by=.(Year,TransectID)]

rx5.sum[,Month:=NULL]
rx5.sum[,Index.m:=NULL]
rx5.sum[order(Year),rx5.sum.lag:= shift(val.mth), by =.(TransectID)]

cover.climate = merge(cover.climate,rx5.sum, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'rx5.sum')

#rx5.year
#rx5.y = indices.monthly.mod[Index.m=="rx5day" & Year>2008,.SD[which.max(val.mth)],by=.(Year,TransectID)]

#rx5.y[,Month:=NULL]
#rx5.y[,Index.m:=NULL]
#rx5.y[order(Year),rx5.y.lag:= shift(val.mth), by =.(TransectID)]

#cover.climate = merge(cover.climate,rx5.y, by=c("Year","TransectID"),all.x = T)
#setnames(cover.climate,"val.mth",'rx5.y')

###d95p

#d95p annual
d95p = indices.monthly.mod[Index.m=="d95p" & Year>2008,sum(val.mth),by=.(Year,TransectID)]

d95p[order(Year),d95p.lag:= shift(V1), by =.(TransectID)]
cover.climate = merge(cover.climate,d95p, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"V1",'d95p')

#d95p.spr
d95p.spr = indices.monthly.mod[Index.m=="d95p" & Year>2008 & (Month =="Apr" | Month == "May" | Month =="Jun"), sum(val.mth),by=.(Year,TransectID)]

d95p.spr[order(Year),d95p.spr.lag:= shift(V1), by =.(TransectID)]
cover.climate = merge(cover.climate,d95p.spr, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"V1",'d95p.spr')

#d95p.win
d95p.win = indices.monthly.mod[Index.m=="d95p" & Year>2008 & (Month =="Dec" | Month == "Jan" | Month == "Feb" | Month=="Mar"), sum(val.mth),by=.(Year,TransectID)]

d95p.win[order(Year),d95p.win.lag:= shift(V1), by =.(TransectID)]
cover.climate = merge(cover.climate,d95p.win, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"V1",'d95p.win')

###ldp

#ldp annual
ldp = indices.monthly.mod[Index.m=="ldp" & Year>2008,.SD[which.max(val.mth)],by=.(Year,TransectID)]

ldp[order(Year),ldp.lag:= shift(val.mth), by =.(TransectID)]
ldp[,Index.m:=NULL]
ldp[,Month:=NULL]
cover.climate = merge(cover.climate,ldp, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'ldp')

#ldp.spr
ldp.spr = indices.monthly.mod[Index.m=="ldp" & Year>2008 & (Month =="Apr" | Month == "May" | Month =="Jun"), .SD[which.max(val.mth)],by=.(Year,TransectID)]

ldp.spr[order(Year),ldp.spr.lag:= shift(val.mth), by =.(TransectID)]
ldp.spr[,Index.m:=NULL]
ldp.spr[,Month:=NULL]
cover.climate = merge(cover.climate,ldp.spr, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'ldp.spr')

#ldp.sum
ldp.sum = indices.monthly.mod[Index.m=="ldp" & Year>2008 & (Month =="Jul" | Month == "Aug"), .SD[which.max(val.mth)],by=.(Year,TransectID)]

ldp.sum[order(Year),ldp.sum.lag:= shift(val.mth), by =.(TransectID)]
ldp.sum[,Index.m:=NULL]
ldp.sum[,Month:=NULL]
cover.climate = merge(cover.climate,ldp.sum, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'ldp.sum')


###lwp

#lwp annual
lwp = indices.monthly.mod[Index.m=="lwp" & Year>2008,.SD[which.max(val.mth)],by=.(Year,TransectID)]

lwp[order(Year),lwp.lag:= shift(val.mth), by =.(TransectID)]
lwp[,Index.m:=NULL]
lwp[,Month:=NULL]
cover.climate = merge(cover.climate,lwp, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'lwp')

#lwp.spr
lwp.spr = indices.monthly.mod[Index.m=="lwp" & Year>2008 & (Month =="Apr" | Month == "May" | Month =="Jun"), .SD[which.max(val.mth)],by=.(Year,TransectID)]

lwp.spr[order(Year),lwp.spr.lag:= shift(val.mth), by =.(TransectID)]
lwp.spr[,Index.m:=NULL]
lwp.spr[,Month:=NULL]
cover.climate = merge(cover.climate,lwp.spr, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'lwp.spr')

#lwp.sum
lwp.sum = indices.monthly.mod[Index.m=="lwp" & Year>2008 & (Month =="Jul" | Month == "Aug"  ), .SD[which.max(val.mth)],by=.(Year,TransectID)]

lwp.sum[order(Year),lwp.sum.lag:= shift(val.mth), by =.(TransectID)]
lwp.sum[,Index.m:=NULL]
lwp.sum[,Month:=NULL]
cover.climate = merge(cover.climate,lwp.sum, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'lwp.sum')



######Drought indices - these are calculated on a rolling window basis, so a 6 month value looks at the current month and the prior 5 months

#No real argument for making 1 month predictions...except for April, which is the beginning of Spring...typically and could be an important precip indicator for invasives

# 1-month on April = early Spring drought (April) - Use regular
# 3-month on June = Spring drought (April, May, June) - Use regular
# 4-month on March = Winter drought (December, Jan, Feb, Mar) - Use Modified
# 3-month on November = Fall drought (Sep, Oct, Nov) - Use Modified
# 2-month on August = Summer drought (July - August) - Use Modified
# 12-month is the past year....maybe less useful than just including seasons individually - Use Modified
# 36-month  is past three years, gives drought overview from the past 3 years since last observation ideally


#HUGH - confirm I am using the correct SPEI value

###SPI...right now I have not calculated SPI at the correct month intervals ^ but the assumption is SPEI might better capture true drought effects 

###SPEI

#1 month april
spei.april = indices.monthly[Index.m=="spei_1_1" & Year>2008 & Month == "Apr",val.mth,by=.(Year,TransectID)]
spei.april[order(Year),spei.april.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,spei.april, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'spei.april')

#2 month August - use modified
spei.sum = indices.monthly.mod[Index.m=="spei_2_1" & Year>2008 & Month == "Aug",val.mth,by=.(Year,TransectID)]
spei.sum[order(Year),spei.sum.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,spei.sum, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'spei.sum')

#3 month spring on Jun - use regular
spei.spr = indices.monthly[Index.m=="spei_3_1" & Year>2008 & Month == "Jun",val.mth,by=.(Year,TransectID)]
spei.spr[order(Year),spei.spr.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,spei.spr, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'spei.spr')

#3 month Fall on Nov - use modified
spei.fall = indices.monthly.mod[Index.m=="spei_3_1" & Year>2008 & Month == "Nov",val.mth,by=.(Year,TransectID)]
spei.fall[order(Year),spei.fall.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,spei.fall, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'spei.fall')

#4 month Winter on March - use Modified
spei.win = indices.monthly.mod[Index.m=="spei_4_1" & Year>2008 & Month == "Nov",val.mth,by=.(Year,TransectID)]
spei.win[order(Year),spei.win.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,spei.win, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'spei.win')

#12 month - use modified
spei.12 = indices.monthly.mod[Index.m=="spei_12_1" & Year>2008 & Month == "Jun",val.mth,by=.(Year,TransectID)]
spei.12[order(Year),spei.12.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,spei.12, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'spei.12')

#36 month - use modified
spei.36 = indices.monthly.mod[Index.m=="spei_36_1" & Year>2008 & Month == "Jun",val.mth,by=.(Year,TransectID)]
spei.36[order(Year),spei.36.lag:= shift(val.mth), by =.(TransectID)]
cover.climate = merge(cover.climate,spei.36, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.mth",'spei.36')


########The rest of the indices are yearly

#Annual indices should be the modified versions to avoid considering weather data that occurs post monitoring

###FD
#fd = indices.yearly[Index.y=='fd' & Year>2008,.(Year,TransectID,val.yr)]
#fd[order(Year),fd.lag:= shift(val.yr), by =.(TransectID)]
#cover.climate = merge(cover.climate,fd, by=c("Year","TransectID"),all.x = T)
#setnames(cover.climate,"val.yr",'fd')

###GSL - use modified
gsl = indices.yearly[Index.y=='gsl.mod' & Year>2008,.(Year,TransectID,val.yr)]
gsl[order(Year),gsl.lag:= shift(val.yr), by =.(TransectID)]
cover.climate = merge(cover.climate,gsl, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.yr",'gsl')

###GSL-start
gsl.start = indices.yearly[Index.y=='ogs6' & Year>2008,.(Year,TransectID,val.yr)]
gsl.start[order(Year),gsl.start.lag:= shift(val.yr), by =.(TransectID)]
cover.climate = merge(cover.climate,gsl.start, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.yr",'gsl.start')

### WSDI - use modified
wsdi = indices.yearly[Index.y=='wsdi.mod' & Year>2008,.(Year,TransectID,val.yr)]
wsdi[order(Year),wsdi.lag:= shift(val.yr), by =.(TransectID)]
cover.climate = merge(cover.climate,wsdi, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.yr",'wsdi')

### CSDI - use modified
#csdi = indices.yearly[Index.y=='csdi.mod' & Year>2008,.(Year,TransectID,val.yr)]
#csdi[order(Year),csdi.lag:= shift(val.yr), by =.(TransectID)]
#cover.climate = merge(cover.climate,csdi, by=c("Year","TransectID"),all.x = T)
#setnames(cover.climate,"val.yr",'csdi')

### SDII - use modified
sdii = indices.yearly[Index.y=='sdii.mod' & Year>2008,.(Year,TransectID,val.yr)]
sdii[order(Year),sdii.lag:= shift(val.yr), by =.(TransectID)]
cover.climate = merge(cover.climate,sdii, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.yr",'sdii')

### CDD - use modified
cdd = indices.yearly[Index.y=='cdd.mod' & Year>2008,.(Year,TransectID,val.yr)]
cdd[order(Year),cdd.lag:= shift(val.yr), by =.(TransectID)]
cover.climate = merge(cover.climate,cdd, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.yr",'cdd')

### CWD - use modified
cwd = indices.yearly[Index.y=='cwd.mod' & Year>2008,.(Year,TransectID,val.yr)]
cwd[order(Year),cwd.lag:= shift(val.yr), by =.(TransectID)]
cover.climate = merge(cover.climate,cwd, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.yr",'cwd')

### PCI - use modified
pci = indices.yearly[Index.y=='pci.mod' & Year>2008,.(Year,TransectID,val.yr)]
pci[order(Year),pci.lag:= shift(val.yr), by =.(TransectID)]
cover.climate = merge(cover.climate,pci, by=c("Year","TransectID"),all.x = T)
setnames(cover.climate,"val.yr",'pci')

```

#Averages and Totals

In addition to these indices we may want to consider total rainfall, and average temps. 

Precipitation totals:
pcpt.yr 
pcpt.spr 
pcpt.win 
pcpt.sum
pcpt.fall

Temperature averages:
temp.yr 
temp.spr 
temp.win 
temp.sum 
temp.fall



``` {r averages-totals}
#Temperature and precip totals/average are already in the correct format, just merge them in. Easiest to convert to wide format

selection = indices.yearly[Year>2008 & (Index.y =="pcpt.yr" |Index.y =="pcpt.spr" |Index.y =="pcpt.win" |Index.y =="pcpt.sum" |Index.y =="pcpt.fall" |Index.y =="temp.yr" |Index.y =="temp.spr" |Index.y =="temp.win" |Index.y =="temp.sum" |Index.y =="temp.fall"),]

selection.wide = dcast(selection,formula =  TransectID + Year ~ Index.y, value.var = 'val.yr')

cover.climate = merge(cover.climate,selection.wide, by = c("TransectID", "Year"))



```

```{r Export}
setwd("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis")

library(data.table)

#Save indices files
fwrite(indices.yearly,"/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis/Indices_Yearly.csv")

fwrite(indices.monthly,"/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis/Indices_Monthly.csv")

#Save cover file with climate data
fwrite(cover.climate,"/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Prairie Analysis/Prairie-Analysis/cover_climate.csv")


```



