---
title: "Climate Variable Calcs"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install daymetr package
# https://cran.r-project.org/web/packages/daymetr/vignettes/daymetr-vignette.html
# install.packages("daymetr")
library(daymetr)
library(data.table)
library(stringr)
library(ClimInd)
library(climdex.pcic)
library(beepr)

#365, 399, 400, 425, 426, 427
#Look at US phenology network for extenced spring

#Load in csv of transect locations
tran.loc = as.data.table(read.csv("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Transect_weather_v2.csv"))

# code is ....partially run
#download_daymet_batch(file_location = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Transect_weather_v3.csv", start = 1980, end = 2018, internal = FALSE, path = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Daymet_2/")



#modify each transect file into one climdex formatted data file per transect

#run climdex input.raw

#check extended spring indices

#use Climind for a hanful of indices...makes sure gsl

#Create list of weather data files by transect
daymet.raw = list.files("/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Daymet_2",full.names = T) 
daymet.raw = daymet.raw[!daymet.raw %in% "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Daymet_2/Icon\r"]

```

#Run all annual indices based on typical year definition. Create a loop to load and modify transect weather data csv then calculate each index and output files into correct bin

Annual indices to run:
WSDI
CDD
CWD
SDII
GSL

Monthly indices to run:
Txx
TX90P
Rx5day
Rx1day


#Climdex indices to run:

First run everything with a normal time step
```{r Normal-climdex}

#Loop CLimdex index calc
#I'm also curious to scrape elevation data for these transects, go ahead and create a data table within this loop

gsl = c()
cdd = c()
wsdi = c()
cwd = c()
sdii = c()

tx90p = c()
rx5day = c()
rx1day = c()
txx = c()

id = c()

elevation.dat = matrix(data = NA, ncol = 2)

for (i  in 1:length(daymet.raw)) {
  tran.weather.dt = as.data.table(read.csv(daymet.raw[[i]],header = T, skip = 7))
  #Properly convert date for climdex input
  tmax.dates = as.PCICt(do.call(paste,tran.weather.dt[,.(year,yday)]),format = "%Y %j", cal = "gregorian")
  tmin.dates = tmax.dates
  prec.dates = tmax.dates
  
  #Create climdexInput object
  ci = climdexInput.raw(tmax = tran.weather.dt$tmax..deg.c.,tmin = tran.weather.dt$tmin..deg.c.,prec = tran.weather.dt$prcp..mm.day., 
                        tmax.dates = tmax.dates, tmin.dates = tmin.dates, prec.dates = prec.dates, 
                        base.range = c(tran.weather.dt[,min(year)],tran.weather.dt[,max(year)]),n = 5, northern.hemisphere = T, 
                        temp.qtiles =  c(0.1, 0.9), prec.qtiles = c(0.95, 0.99), max.missing.days = c(annual = 5, monthly = 3),
                        min.base.data.fraction.present = 0.1)
  
  TransectID = as.numeric(str_sub(basename(daymet.raw[[i]]),end = -15))
  
  #Grab elevation data
  elevation = as.numeric(str_sub(str_sub(as.character(read.csv(daymet.raw[[i]],header = T, skip = 2,nrows = 1)[1,1]),start = 12),end = -8))
  
  elevation.dat = rbind(elevation.dat, c(TransectID, elevation))
  
  #Yearly Climdex Indices
  gsl.dat = climdex.gsl(ci)
  gsl.dat = c(TransectID,gsl.dat)
  gsl = c(gsl, gsl.dat)
  
  cdd.dat = climdex.cdd(ci)
  cdd.dat = c(TransectID,cdd.dat)
  cdd = c(cdd, cdd.dat)
  
  wsdi.dat = climdex.wsdi(ci)
  wsdi.dat = c(TransectID,wsdi.dat)
  wsdi = c(wsdi, wsdi.dat)
  
  cwd.dat = climdex.cwd(ci)
  cwd.dat = c(TransectID,cwd.dat)
  cwd = c(cwd, cwd.dat)
  
  sdii.dat = climdex.sdii(ci)
  sdii.dat = c(TransectID,sdii.dat)
  sdii = c(sdii, sdii.dat)

  id = c(id, TransectID)
  
  #Monthly Climdex Indices
  txx.dat = climdex.txx(ci, freq = "monthly")
  txx = c(txx, txx.dat)
  
  rx1day.dat = climdex.rx1day(ci, freq = "monthly")
  rx1day = c(rx1day, rx1day.dat)
  
  rx5day.dat = climdex.rx5day(ci, freq = "monthly")
  rx5day = c(rx5day, rx5day.dat)
  
  tx90p.dat = climdex.tx90p(ci, freq = "monthly")
  tx90p = c(tx90p, tx90p.dat)

}

beep(sound = 2)
  

gsl = matrix(gsl, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
cdd = matrix(cdd, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
cwd = matrix(cwd, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
sdii = matrix(sdii, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
wsdi = matrix(wsdi, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)

colnames(gsl) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(wsdi) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(cdd) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(cwd) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(sdii) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])

#Save output csv
write.csv(x = gsl, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/gsl.csv")
write.csv(x = cdd, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/cdd.csv")
write.csv(x = cwd, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/cwd.csv")
write.csv(x = sdii, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/sdii.csv")
write.csv(x = wsdi, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/wsdi.csv")

#Now organize the monthly indices
txx.1 = as.data.table(matrix(txx, ncol = 12, byrow = F))
txx.1[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
txx.1[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(txx.1) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

tx90p.1 = as.data.table(matrix(tx90p, ncol = 12, byrow = F))
tx90p.1[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
tx90p.1[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(tx90p.1) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

rx1day.1 = as.data.table(matrix(rx1day, ncol = 12, byrow = F))
rx1day.1[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
rx1day.1[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(rx1day.1) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

rx5day.1 = as.data.table(matrix(rx5day, ncol = 12, byrow = F))
rx5day.1[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
rx5day.1[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(rx5day.1) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

#Save output csv
write.csv(x = txx.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/txx.csv")
write.csv(x = tx90p.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/tx90p.csv")
write.csv(x = rx5day.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/rx5day.csv")
write.csv(x = rx1day.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/rx1day.csv")

```

Now run annual indices with modified dates. There are no monitoring actions before July 1, lets use this as the cut-off day. Year day = 182

```{r modified-climdex}

#Loop CLimdex index calc
#I'm also curious to scrape elevation data for these transects, go ahead and create a data table within this loop

gsl.mod = c()
cdd.mod = c()
wsdi.mod = c()
cwd.mod = c()
sdii.mod = c()

id = c()


for (i in 1:length(daymet.raw)) {
  tran.weather.dt = as.data.table(read.csv(daymet.raw[[i]],header = T, skip = 7))
  
  #Modify the data such they year effectively ends after June
  tran.weather.dt[yday>181,year:= year-1]
  
  #Properly convert date for climdex input
  tmax.dates = as.PCICt(do.call(paste,tran.weather.dt[,.(year,yday)]),format = "%Y %j", cal = "gregorian")
  tmin.dates = tmax.dates
  prec.dates = tmax.dates
  
  #Create climdexInput object
  ci = climdexInput.raw(tmax = tran.weather.dt$tmax..deg.c.,tmin = tran.weather.dt$tmin..deg.c.,prec = tran.weather.dt$prcp..mm.day., 
                        tmax.dates = tmax.dates, tmin.dates = tmin.dates, prec.dates = prec.dates, 
                        base.range = c(tran.weather.dt[,min(year)],tran.weather.dt[,max(year)]),n = 5, northern.hemisphere = T, 
                        temp.qtiles =  c(0.1, 0.9), prec.qtiles = c(0.95, 0.99), max.missing.days = c(annual = 200, monthly = 3),
                        min.base.data.fraction.present = 0.1)
  
  TransectID = as.numeric(str_sub(basename(daymet.raw[[i]]),end = -15))
  
  #Yearly Climdex Indices
  gsl.dat = climdex.gsl(ci)
  gsl.dat = c(TransectID,gsl.dat)
  gsl.mod = c(gsl.mod, gsl.dat)
  
  cdd.dat = climdex.cdd(ci)
  cdd.dat = c(TransectID,cdd.dat)
  cdd.mod = c(cdd.mod, cdd.dat)
  
  wsdi.dat = climdex.wsdi(ci)
  wsdi.dat = c(TransectID,wsdi.dat)
  wsdi.mod = c(wsdi.mod, wsdi.dat)
  
  cwd.dat = climdex.cwd(ci)
  cwd.dat = c(TransectID,cwd.dat)
  cwd.mod = c(cwd.mod, cwd.dat)
  
  sdii.dat = climdex.sdii(ci)
  sdii.dat = c(TransectID,sdii.dat)
  sdii.mod = c(sdii.mod, sdii.dat)

  id = c(id, TransectID)

}

beep(sound=2)

gsl.mod.1 = matrix(gsl.mod, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
cdd.mod.1 = matrix(cdd.mod, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
cwd.mod.1 = matrix(cwd.mod, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
sdii.mod.1 = matrix(sdii.mod, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
wsdi.mod.1 = matrix(wsdi.mod, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)

colnames(gsl.mod.1) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(wsdi.mod.1) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(cdd.mod.1) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(cwd.mod.1) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(sdii.mod.1) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])

#Save output csv
write.csv(x = gsl.mod.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/gsl.mod.csv")
write.csv(x = cdd.mod.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/cdd.mod.csv")
write.csv(x = cwd.mod.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/cwd.mod.csv")
write.csv(x = sdii.mod.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/sdii.mod.csv")
write.csv(x = wsdi.mod.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/wsdi.mod.csv")



```

#Now run the extra indices from ClimInd
Adding a few new indices to consider:

pci - Index to evaluate precipitation heterogeneity at a monthly scale. Values <10 (uniform monthly rainfall distribution); values 11-15 (moderate concentration of precipitation); values 16-20 (irregular distribution); and >20 ((high precipitation concentration)

spei - 1, 3, 12 month

spi - 1, 3, 12



```{r Normal-climInd_SPEI}

#ClimInd uses slightly different inputs
#Evaluate structure of example data set 
#summary(data(data("data_all"))) #All of the data vectors use the dates as the name


d95p = data.frame(Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA) #make this monthly, Days with precipitation > 95p
ldp = data.frame(Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA) #Longest dry period, could make this monthly but would be missing dry period which span different months
lwp = data.frame(Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA) #longest wet period
ogs6 = c() #onset of growing season 6 days, matches the other gsl index
pci = c()

spi1 =  data.frame(Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA)
spi3 = data.frame(Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA)
spi6 = data.frame(Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA)
spi12 = data.frame(Jan = 1,Feb = 1,Mar =1, Apr=1,May=1,Jun=1,Jul=1,Aug=1,Sep=1,Oct=1,Nov=1,Dec=1,Year = 1,TransectID=NA)

id = c()

for (i in 1:length(daymet.raw)) {


tran.weather.dt = as.data.table(read.csv(daymet.raw[[i]],header = T, skip = 7))

rr = tran.weather.dt$prcp..mm.day.
tg = apply(tran.weather.dt[,.(tmax..deg.c.,tmin..deg.c.)],MARGIN = 1, FUN = mean ) #Daily mean
tn = tran.weather.dt$tmin..deg.c.
tx = tran.weather.dt$tmin..deg.c.
radiaton_w = tran.weather.dt$srad..W.m.2. #shortwave radiation
vapour = tran.weather.dt$vp..Pa. #Water vapor pressure in pascals. Daily average partial pressure of water vapor.

#Convert dates and format for ClimInd
tmax.dates = as.PCICt(do.call(paste,tran.weather.dt[,.(year,yday)]),format = "%Y %j", cal = "gregorian")
dates  =  noquote(format(tmax.dates,"%m/%d/%y"))

names(rr) = dates
names(tg) = dates
names(tn) = dates
names(tx) = dates
names(radiaton_w) = dates
names(vapour) = dates

TransectID = as.numeric(str_sub(basename(daymet.raw[[i]]),end = -15))

id = c(id, TransectID)

#Now loop through indices
d95p.dat = d95p(rr, time.scale = 'month')
d95p.dat = cbind(d95p.dat,Year=as.numeric(row.names(d95p.dat)))
d95p.dat = cbind(d95p.dat, TransectID = rep(TransectID, times = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])))
d95p = rbind(d95p, d95p.dat)

ldp.dat = ldp(rr, time.scale = 'month')
ldp.dat = cbind(ldp.dat,Year=as.numeric(row.names(ldp.dat)))
ldp.dat = cbind(ldp.dat, TransectID = rep(TransectID, times = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])))
ldp = rbind(ldp, ldp.dat)

lwp.dat = lwp(rr, time.scale = 'month')
lwp.dat = cbind(lwp.dat,Year=as.numeric(row.names(lwp.dat)))
lwp.dat = cbind(lwp.dat, TransectID = rep(TransectID, times = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])))
lwp = rbind(lwp, lwp.dat)

spi1.dat = spi1(rr)
spi1.dat = cbind(spi1.dat,Year=as.numeric(row.names(spi1.dat)))
spi1.dat = cbind(spi1.dat, TransectID = rep(TransectID, times = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])))
spi1 = rbind(spi1, spi1.dat)

spi3.dat = spi3(rr)
spi3.dat = cbind(spi3.dat,Year=as.numeric(row.names(spi3.dat)))
spi3.dat = cbind(spi3.dat, TransectID = rep(TransectID, times = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])))
spi3 = rbind(spi3, spi3.dat)

spi6.dat = spi6(rr)
spi6.dat = cbind(spi6.dat,Year=as.numeric(row.names(spi6.dat)))
spi6.dat = cbind(spi6.dat, TransectID = rep(TransectID, times = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])))
spi6 = rbind(spi6, spi6.dat)

spi12.dat = spi12(rr)
spi12.dat = cbind(spi12.dat,Year=as.numeric(row.names(spi12.dat)))
spi12.dat = cbind(spi12.dat, TransectID = rep(TransectID, times = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])))
spi12 = rbind(spi12, spi12.dat)

ogs6.dat = ogs6(tg)
ogs6.dat = c(TransectID,ogs6.dat)
ogs6 = c(ogs6, ogs6.dat)

pci.dat = pci(rr)
pci.dat = c(TransectID,pci.dat)
pci = c(pci, pci.dat)

}
beep(sound=2)

#Modify yearly index outputs and save as csv
ogs6.1 = matrix(ogs6, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)
pci.1 = matrix(pci, ncol = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]) + 1, byrow = T)

colnames(ogs6.1) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])
colnames(pci.1) = c("TransectID",tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)])

#Save output csv
write.csv(x = ogs6.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/ogs6.csv")
write.csv(x = pci.1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Yearly/pci.csv")

#Remove top filler row from montly data
d95p = d95p[-1,]
ldp = ldp[-1,]
lwp = lwp[-1,]
spi1 = spi1[-1,]
spi3 = spi3[-1,]
spi6 = spi6[-1,]
spi12 = spi12[-1,]

#Save monthly outputs
write.csv(x = d95p, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/d95p.csv")
write.csv(x = ldp, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/ldp.csv")
write.csv(x = lwp, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/lwp.csv")
write.csv(x = spi1, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spi1.csv")
write.csv(x = spi3, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spi3.csv")
write.csv(x = spi6, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spi6.csv")
write.csv(x = spi12, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spi12.csv")

```

#SPEI
This will take some additional work since I don't have ETO precalculated, use the SPEI package

To calculate Potential Evapotranspiration I will use the Hargreaves method, which allows me to incorporare radiation in without needing wind speed like in the penman method. 

```{r SPEI}
#First calculate PET
#This uses Tmin, Tmax, radiation, latitude, and precip...will need to convert radiation values (monthly mean daily)

spei.1.1 = c()
spei.1.2 = c()

spei.3.1 = c()
spei.3.2 = c()

spei.6.1 = c()
spei.6.2 = c()

spei.12.1 = c()
spei.12.2 = c()

id = c()


for (i in 1:length(daymet.raw)) {

  tran.weather.dt = as.data.table(read.csv(daymet.raw[[i]],header = T, skip = 7))

  #Downscale daily data into monthly averages, requires adding month value to data

tran.weather.dt[,month:= month(as.PCICt(do.call(paste,tran.weather.dt[,.(year,yday)]),format = "%Y %j", cal = "gregorian",label = FALSE))]

  Tmin = ts(tran.weather.dt[,mean(tmin..deg.c.), by = .(month,year)][,V1],start = 1980, frequency = 12)
  Tmax = ts(tran.weather.dt[,mean(tmax..deg.c.), by = .(month,year)][,V1],start = 1980, frequency = 12)
  Ra = ts(tran.weather.dt[,mean((srad..W.m.2.*dayl..s.)/1000000), by = .(month,year)][,V1],start = 1980, frequency = 12) #also converts to daily radiation (MJ/m2/day)
  Pre = ts(tran.weather.dt[,sum(prcp..mm.day.), by = .(month,year)][,V1],start = 1980, frequency = 12)
  PRCP = ts(tran.weather.dt[,mean(prcp..mm.day.), by = .(month,year)][,V1],start = 1980, frequency = 12)
  TransectID = as.numeric(str_sub(basename(daymet.raw[[i]]),end = -15))

  lat = tran.loc[Site == TransectID,Latitude]

  #PET.1 = hargreaves(Tmin, Tmax, Ra, lat, Pre)
  PET.2 = hargreaves(Tmin, Tmax, lat = lat)

  #I don't know which method is better...run them all for now
  #bal.1 = Pre - PET.1
  bal.2 = Pre - PET.2


  #spei.1.1 = c(spei.1.1,c(spei(bal.1, 1)$fitted)) #calculate 1 month spei
  spei.1.2 = c(spei.1.2,c(spei(bal.2, 1)$fitted)) 

  #spei.3.1 = c(spei.3.1,c(spei(bal.1, 3)$fitted)) 
  #spei.3.2 = c(spei.3.2,c(spei(bal.2, 3)$fitted)) 

  #spei.6.1 = c(spei.6.1,c(spei(bal.1, 6)$fitted)) 
  #spei.6.2 = c(spei.6.2,c(spei(bal.2, 6)$fitted)) 

  #spei.12.1 = c(spei.12.1,c(spei(bal.1, 12)$fitted)) 
  #spei.12.2 = c(spei.12.2,c(spei(bal.2, 12)$fitted)) 

  id = c(id, TransectID)

}

beep(sound = 2)

#coerce spei vectors in matrices and add identifying columns
spei.1.1.m  = as.data.table(matrix(spei.1.1, ncol = 12, byrow = T))
spei.1.1.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.1.1.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.1.1.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

spei.1.2.m  = as.data.table(matrix(spei.1.2, ncol = 12, byrow = T))
spei.1.2.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.1.2.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.1.2.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

spei.3.1.m  = as.data.table(matrix(spei.3.1, ncol = 12, byrow = T))
spei.3.1.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.3.1.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.3.1.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

spei.3.2.m  = as.data.table(matrix(spei.3.2, ncol = 12, byrow = T))
spei.3.2.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.3.2.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.3.2.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

spei.6.1.m  = as.data.table(matrix(spei.6.1, ncol = 12, byrow = T))
spei.6.1.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.6.1.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.6.1.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

spei.6.2.m  = as.data.table(matrix(spei.6.2, ncol = 12, byrow = T))
spei.6.2.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.6.2.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.6.2.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

spei.12.1.m  = as.data.table(matrix(spei.12.1, ncol = 12, byrow = T))
spei.12.1.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.12.1.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.12.1.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

spei.12.2.m  = as.data.table(matrix(spei.12.2, ncol = 12, byrow = T))
spei.12.2.m[,Year:= rep(c(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]),times = length(id))]
spei.12.2.m[,TransectID:= rep(id,each = length(tran.weather.dt[,min(year)]:tran.weather.dt[,max(year)]))]
colnames(spei.12.2.m) = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec','Year','TransectID')

#save csv output
write.csv(x = spei.1.1.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_1_1.csv")
write.csv(x = spei.1.2.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_1_2.csv")
write.csv(x = spei.3.1.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_3_1.csv")
write.csv(x = spei.3.2.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_3_2.csv")
write.csv(x = spei.6.1.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_6_1.csv")
write.csv(x = spei.6.2.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_6_2.csv")
write.csv(x = spei.12.1.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_12_1.csv")
write.csv(x = spei.12.2.m, file = "/Volumes/GoogleDrive/My Drive/Thesis/Prairies/Data And Analysis/Climate/GMT/Indices-Monthly/spei_12_2.csv")


```
